### 2024-05-28 README.Rmd
More edits!!

## Background

@@

## The [mission](https://president.mcmaster.ca/mission-vision/)

McMaster's mission is "the discovery, communication and preservation of knowledge," accomplished with a commitment to "creativity, innovation and excellence". These aspirations must respond to some underlying values, and indeed, the university explicitly values "integrity, quality, inclusiveness and teamwork". Ultimately, the university aims to "inspire critical thinking, personal growth, and a passion for lifelong learning" and to serve "the social, cultural, and economic needs of our community and our society."

@@

## What do AIs do?

@@

## A note on the terms of the debate

@@

## But really, what do AIs do?

@@

## AI in teaching and learning

@@

## What can GenAI tools do for students?

@@

## What do GenAI tools do to students?

@@

Mind you, these, and numerous other [hapless results](https://www.nbcnews.com/tech/tech-news/google-ai-im-feeling-depressed-cheese-not-sticking-to-pizza-error-rcna153301) (as of May, 2023) are produce by AI _after_ Google has already spent tens of billions of dollars on the technology [@vynck_after_2024]. To what extent can more money improve AI? Probably not by much, since the key constraint is the amount of data available for training.

@@

## What do GenAIs do for the mega-corporations pushing these tools?

@@

## What do GenAIs to to everyone else?

@@

Above, the ways in which GenAI is exploitative of people all over the world, but particularly in the Global South were highlighted. But AI companies do not spare the rich either. OpenAI's Sam Altman, a self-avowed admirer of the movie _Her_, featuring Scarlet Johansson as an AI voice assistant, reportedly tried to hire Ms. Johannson to be the voice of ChatGPT 4.0 [@kastrenakes_scarlett_2024]. After Ms. Johannson decline, OpenAI went and used a mimicry of [her voice](@as-an-aside) anyways, which was retired after Ms. Johansson threatened legal action^[Altman's denials that OpenAI had not used Ms. Johansson's voice sounded hollow, after he [posted](https://x.com/sama/status/1790075827666796666) on social media the word "her" at the time the new version of ChatGPT was announced.] [@allyn_scarlett_2024]. Alas, very few have the social standing of Scarlett Johansson to challenge OpenAI. Most people in the world lack that power.

@@

@@

### Replication of unjust structures of power

When using a Generative AI we must ask whose perspectives that tool represents.

@@

ChatGPT was an awesome feat of marketing [@cowen_chatgpt_2023], but much coverage of the relevant issues in the media has been so credulous it has stretched credulity past Pluto [@burneko_if_2024]^[Bob Woodward said "journalism is not stenography" [@hiskes_woodward_2007], but it is difficult to see what else writing like Roose's [-@roose_is_2024] could be called.]. There are now calls to press pause on the hype [@angwin_press_2024] but these are still drown by the credulous coverage.

>>>>

---

## Contents

- [Background](#background)  
- [The university's mission](#the-mission)  
- [What do AIs do?](#what-do-ais-do)  
- [A note on the terms of the conversation](#terms-of-debate)  
- [But really, what do AIs do?](#really-what-do-they-do)  
- [AI in teaching and learning](#ai-in-teaching-and-learning)  
- [What can GenAI tools do for students?](#what-can-ais-do-for-students)  
- [What do GenAI tools do to students?](#what-can-ais-do-to-students)  
- [What do GenAIs do for the mega-corporations pushing these tools?](#what-do-ais-to-corporations)  
- [What do GenAIs to to everyone else?](#what-do-ais-do-to-everyone-else)

---

## Background {#background}

@@

## The university's mission {#the-mission}

McMaster's [mission](https://president.mcmaster.ca/mission-vision/) is "the discovery, communication and preservation of knowledge," accomplished with a commitment to "creativity, innovation and excellence". These aspirations must respond to some underlying values, and indeed, the university explicitly values "integrity, quality, inclusiveness and teamwork". Ultimately, the university aims to "inspire critical thinking, personal growth, and a passion for lifelong learning" and to serve "the social, cultural, and economic needs of our community and our society."

@@

## What do AIs do?{#what-do-ais-do}

@@

## A note on the terms of the conversation {#terms-of-debate}

@@

## But really, what do AIs do? {#really-what-do-they-do}

@@

## AI in teaching and learning {#ai-in-teaching-and-learning}

@@

## What can GenAI tools do for students?{#what-can-ais-do-for-students}

@@

## What do GenAI tools do to students? {#what-can-ais-do-to-students}

@@

Mind you, these, and numerous other hapless results^[For instance, ; or returning misinformation about depresion [@tenbarge_glue_2024]] (as of May, 2024) are produced by an AI that Google has already spent tens of billions of dollars developing [@vynck_after_2024]. To what extent can more money improve AI? Probably not by much, since the key constraint is the amount of data available for training, and as numerous examples

a search for “how to rotate text in ms paint,” referring to the Microsoft Paint application. The top Google search result, viewed by NBC News, directs the reader to a sarcastic Reddit comment that says to press the “Flubblegorp” key on your keyboard. This key does not exist.

@@

## What do GenAIs do for the mega-corporations pushing these tools? {#what-do-ais-to-corporations}

@@

## What do GenAIs to to everyone else? {#what-do-ais-do-to-everyone-else}

@@

ChatGPT 4.o was announced on May 13, 2024, with a flashy display of the model's new capabilities, which included the ability to respond vocally to prompts, and to identify images via a mobile device's camera. 4.o's voice, in particular, caused some titillation in the pants of even its creators [@roose_is_2024]. The voice was modeled after the digital personal assistant of the movie _Her_, played by Scarlett Johannson. This film (to approximately no one's surprise) is, by his own account, one of OpenAI's Altman's favorites. The background story of 4.o's voice is somewhat sordid: Altman reportedly tried to hire Ms. Johannson to be the voice of ChatGPT 4.o [@kastrenakes_scarlett_2024], but when Ms. Johannson declined, OpenAI went and used a mimicry of [her voice](@as-an-aside) anyways. This voice was retired (probably to the profound sadness of many an AI engineer) after Ms. Johansson threatened legal action^[Altman's denials that OpenAI had not used Ms. Johansson's voice sounded hollow, after he [posted](https://x.com/sama/status/1790075827666796666) on social media the word "her" at the time the new version of ChatGPT was announced.] [@allyn_scarlett_2024].

Alas, very few have the social standing of Scarlett Johansson to challenge OpenAI. Most people in the world lack that power.

The ways in which GenAI is exploitative of people all over the world, but particularly in the Global South were highlighted. But AI companies do not spare the rich either.

@@

@roberts_commercial_2016 Commercial {Content} {Moderation}: {Digital} {Laborers}' {Dirty} {Work}

@@

### Reproduction of unjust structures of power

When using a Generative AI we must ask whose perspectives the tool will tend to reproduce and amplify.

Training data, much like technology, are not politically neutral, and several processes contribute to reproduce malignant biases, A way to explain this is by means of the concept of _platformed racism_ [@matamoros-fernandez_platformed_2017]. Platformed racism is the process whereby the functionalities of platforms are exploited to create and spread racist ideology (liking and sharing), often in a grass-roots fashion, but more ominously with the assist of bad-faith actors^[Online Russian disinformation has had a wide reach [@treyger_russian_2022] and furthermore is known to exploit "all kinds of separatism and ethnic, social and racial conflicts, actively supporting all dissident movements--extremist, racist, and sectarian groups" to destabilize internal U.S. politics [@us_department_of_state_gec_2024, p. 51]. Given how cagey corporations are about the data they use, it is anyone's guess how much dis- and misinformation is in the datasets used to train their AIs (the Unfathomable Data problem discussed by @bender_dangers_2021).]. The second aspect of platformed racism is the implicit endorsement of racist ideology through vague standards and [arbitrary moderation practices](https://www.bbc.com/news/technology-58058428) that exclude some views while allowing others to thrive [@myers_west_censored_2018; @hawkins_race_2023].

Search engines were already known to reinforce racism [@noble_algorithms_2018]; given the above, it is too mych to expecet that AI-augmented search engines do better^[A report by the Center for Countering Digital Hate found that Google's Bard generated wrong and/or hateful outputs 78 out of 100 tests [@center_for_countering_digital_hate_googles_2023]. In addition to being misleading to the unadvised, it is also a stupendous tool ripe for abuse by the advised with an agenda.]. In another wretched incident, Google's AI overview stated that Barack Obama was the first Muslim president of the US, a regrettable piece of misinformation that fails to surprise given the years-long right-wing disinformation campaign to otherize Obama [@tenbarge_glue_2024]. Mishaps like this, and others, are dismissed by Google as attempts to "to trip up the technology with uncommon questions" [@tenbarge_glue_2024], which only reinforces the notion that AIs only are useful when [regressing to the mean](#really-what-do-they-do). Responses like fiddling with the model (fine tune it) to provide more palatable outputs (like in the case of **FIND EXAMPLE**) are not more reassuring. Who is to say what a handful of unaccountable corporations will find palatable tomorrow, let alone in the face of powerful authoritarian regimes?

```{r google-obama, out.width="50%", fig.cap="\\label{fig:google-ai-overview-of-obama}Google's AI overview of Muslim US presidents"}
google_plot <- ggplot() +
  labs(caption = "Source: https://www.nbcnews.com/tech/tech-news/google-ai-im-feeling-depressed-cheese-not-sticking-to-pizza-error-rcna153301") +
  theme_void()

ggdraw() + 
  draw_plot(google_plot) +
  draw_image(image = "images/google-obama.png")
```

@@

ChatGPT was an awesome feat of marketing [@cowen_chatgpt_2023], but much coverage of the relevant issues in the media has been so credulous it has stretched credulity past Pluto [@burneko_if_2024]^[Bob Woodward said "journalism is not stenography" [@hiskes_woodward_2007], but it is difficult to see what else writing like Roose's [-@roose_is_2024] could be called, besides hagiography.]. There are now calls to press pause on the hype [@angwin_press_2024] but these are still drown by the credulous coverage.
